Fixing Vulnerabilities Potentially Hinders Maintainability

### 2021-02-07 paper/emse20.tex
adding lessons learned

The \emph{Write Code Once} guideline results show that duplicated 
code increased in $171$ ($17.7\%$) of the patches. Software systems 
typically have $9\%$-$17\%$ of cloned code~\cite{5773403}. Previous 
work showed a correlation between code smells and code
duplication~\cite{7476787} which may also be reflected in the 
\emph{Write Clean Code} guideline results. BCH reported new code 
smells for $145$ ($15.0\%$) patches which according to previous work 
may harm software security since some of these new code smells may 
be new vulnerabilities~\cite{8819456}. Thus, while developers 
tend to ignore patch techniques such as the \emph{Replace Method 
with Method Object} technique, they may also be introducing new 
software vulnerabilities and, consequently, harm the market value 
and economy of companies~\cite{4267025}.

@@

>>>>

The \emph{Write Code Once} guideline results show that duplicated 
code increased in $17.7\%$ ($171$/$969$) of the patches. Software systems 
typically have $9\%$-$17\%$ of cloned code~\cite{5773403}. Previous 
work showed a correlation between code smells and code
duplication~\cite{7476787} which may also be reflected in the 
\emph{Write Clean Code} guideline results. BCH reported new code 
smells for $15.0\%$ ($145$/$969$) of the patches which according to previous work 
may be the source of new software vulnerabilities~\cite{8819456,10.1145/3133956.3134072}
capable of harming the market value and economy of companies~\cite{4267025}.
Developers should never reuse code by copying and pasting 
existing code fragments. Instead, they should create a method and call 
it every time needed. The \emph{Extract Method} refactoring 
technique solves many duplication problems. This makes spoting and 
solving the issue faster because you only need to fix the method used
instead of locating and fixing the issue multiple times.
Clone detection tools can also help on locating the issues.

@@

\section{Lessons Learned}\label{sec:lessons}

In this section, we summarize the lessons learned and 
best practices developers should follow to perform 
quality security patches and avoid hindering 
software maintainability.

\textit{\textbf{Best Practices:}} Developers are not 
paying attention to some quality aspects of their solutions/patches, 
as seen in Figure~\ref{fig:guidelines}, ending up harming 
software maintainability. We argue that developers
should design and implement solutions that respect the 
limit bounds of branch points and function/module sizes that are 
recommended by best practices to avoid increasing the 
size, complexity and dependencies of their patches.
Still on respecting bound limits, developers 
should also consider the limit of $4$ parameters per function---to 
keep unit interfaces small and easy to use and understand.
Patterns such as \emph{Introduce Parameter Object} are useful 
to send information to a new function/class through an object
and keep the number of parameters small and the information well-organized.

Security patches also have a negative impact on the maintenance of 
software architecture. Maintain the components independence and balance 
is important to make it easier to find the source code that 
developers want to patch/improve and to better understand
how the high-level componenet interact with others. Applying 
\emph{encapsulation} to hide implementation details and make the system 
more modular is a step forward to not hinder software architecture maintainability.

According to previous research, there is a correlation between 
code duplication and code smells~\cite{7476787}---duplicates are a source 
of regression bugs. BCH reports new code smells
for $15\%$ of the patches under study which supports previous research---$34\%$ of security patches 
introduce new problems~\cite{8819456,10.1145/3133956.3134072}. 
Developers should never reuse code by copying and pasting 
existing code fragments. Instead, they should create a method and call 
it every time needed. The \emph{Extract Method} refactoring 
technique solves many duplication problems. This makes spoting and 
solving the issue faster because you only need to fix one method instead
of multiple vulnerabilities. Clone detection tools such as CPD can help 
on locating duplicates.

\textit{\textbf{Severity:}} Research in the past exhibits proof that developers prioritize
higher impact vulnerabilities~\cite{10.1145/3133956.3134072}.
Our study results show that vulnerabilities of high and
medium severity should be prioritized in software maintainability
by developers.

\textit{\textbf{Programming Language:}} In general, our study 
shows that there is over $35\%$ of patches with negative impact 
on software maintainability per programming language. Developers implementing code in   
\texttt{C/C++}, \texttt{Ruby} and \texttt{PHP} should pay extra care 
to their solutions/patches.

\textit{\textbf{Types of vulnerabilities:}} We tried to understand 
the impact of different types of vulnerabilities on software maintainability.
Overall, all the CWEs studied in more depth present a negative impact
over $30\%$ on software maintainability. 
\emph{Cross-Site Scripting} (CWE-79) and \emph{Improper 
Restriction of Operations within the Bounds of a Memory Buffer} (CWE-119) 
are less prone to have impact on open-source software maintainability.
Developers should pay special attention to \emph{Improper Input Validation} (CWE-20), 
\emph{Information Exposure} (CWE-200), \emph{Missing Release of Memory after Effective Lifetime} 
(CWE-401) and \emph{Path Traversal} (CWE-22). However, 
more research should be performed to better understand 
the impact of each guideline on each CWE.
